HO unification from object language to meta language

### 2024-02-20 paper/main.tex
tex update

In order to reason about unification of the terms of an objet language within a
meta lanauge, we start by formally describing the two languages. We use
meta-programming to accomplish this result. In \cref{code:lang-descr}, we
provide on the left the term representation of the object language. %
% Variables in the OL and ML are represented with integers
In this encoding, the fo\_tm type stands for first order representation of the
object language terms. We model unification variables as integers corresponding
to memory adresses. This memory is represented witht the subst type which is 
list of optional terms. If the cell $i$ is none, then the variable $i$ is not
assigned.

% Why the fo prefix for fo_tm && Uvar in OL have no scope
This encoding is called first order, since variables have no scope. To better
understand this definition, let's take the subterm \coqIn{P x} from the instance
forall\_dec. \coqIn{P} has type \coqIn{A ~$\to$~ Prop} and \coqIn{x} is a name
bound to \coqIn{P}. The application of \coqIn{P} to \coqIn{x} at the OL level,
put this name in the scope of the variable \coqIn{P}. However, at the meta
level, the translation of \coqIn{P x} become \elpiIn{app[uv 1, x]}; \elpiIn{x}
is not in the scope of \elpiIn{uv 1}.

% Uvar in ML have scope
To solve this problem, we give more expressiveness to the OL variables, by
giving an alternative term encoding, where unification variables are the
variables of the meta language. These variables have a scope and therefore
can be of higher order.
per la equal, 
% The subst mappings
In order to represent variable assignation, we declare the two types subst\_fo
and subst coressponding to the substitution mapping of unification variable
for both the OL and the ML.

@@

% We have two different equal functions
Unification over terms is possible thanks to the equal predicate. There exist
two version of this predicate: \eqfo and \eqho. Their implementation is given in
\cref{code:term-equal}. The first tests if two terms are equal in the object
langauge, that is, given a substitution $\theta$, is it true that the two terms
are equal? Note that no variable instanciation is done: unification is not
performed by the \elpiIn{equal} predicate. Therefore, a unification variabile
$i$ equal to another term $t$ if $t$ is a variable with same index $i$, or if
$i$ is assigend to $t'$ in $\theta$ and \elpiIn{equal ~$t'$~ ~$t$~}. Moreover
\eqfo, is able to verify if two terms are equal up to \etabeta. This equality
relation is in fact the equality relation of the object langauge and therefore,
since it accepts \etabeta equality, then \eqfo should take into account these
reductions.

The \eqho predicate test if two terms are equal in the world of the meta
languages. This eaquality function is not capable to understand the \etabeta
reductions of the OL. This is mainly due to the fact that the meta language 
understand \etabeta, but, since the nodes \elpiIn{app} and \elpiIn{lam} do
not respectively represent the internal application and lambda abstraction
of the language, the \etabeta reducation is not possible.

As we can

% Compile FO to HO

>>>>

\def\eqfo{eq\_fo\xspace}

In order to reason about unification of the terms of an objet language within a
meta language, we start by formally describing the two languages. We use
meta-programming to accomplish this result. In \cref{code:lang-descr}, we
provide on the left the typing structure of the terms of the object language
together with the signature of the \eqfo function.

In this encoding, the fo\_tm type stands for the first order representation of
the object language terms. We model unification variables as integers
corresponding to memory addresses. This memory is represented with the subst
type abbreviation which is list of optional terms. If the cell $i$ is none, then
the variable $i$ is not assigned.

% Why the fo prefix for fo_tm && Uvar in OL have no scope
Because of our translation of the terms of the object language into the meta
language we have sketched in section one, a object language term has no scope.
An example of this claim can be given by analyzing deeply the subterm \coqIn{P x} from
the instance forall\_dec. \coqIn{P} is an higher order variables
with type \coqIn{A ~$\to$~ Prop} and
\coqIn{x} is a name bound to \coqIn{P}.
However, at the meta level, the translation of \coqIn{P x} become \elpiIn{app[uv
1, x]}, that is, \elpiIn{uv 1} cannot see \elpiIn{x}.

Since our goal is to work with higher order variables, we need to compile the
term received in entry into a more expressive version. This is possible by using
a second type structure for the terms of the OL. This term definition is given
in the second column of the \cref{code:lang-descr}. The main different here is
that a unification variable (always identified with by an integer) has now a
list of terms representing the scope of the variable. We symbolize the
abstractions of the ML using the constructor \elpiIn{abs} of the type
\elpiIn{assmt}. The constructor \elpiIn{val} is is used to contain terms of the
object language. In this context we do not need to have an application node for
the ML.

@@

% We have two different equal functions
The equality relation over terms under a given substitution mapping is possible 
thanks to the equal predicate. There exist
two version of this predicate: \eqfo and \eqho. Their implementation is given in
\cref{code:term-equal}.

The first tests if two terms are equal in the object language, that is, given a
substitution $\theta$, is it true that the two terms are equal? Note that no
variable instantiation is done: unification is not performed by the
\elpiIn{equal} predicate. Therefore, a unification variable $i$ equals to
another term $t$ if $t$ is a variable with same index $i$, or if $i$ is assigned
to $t'$ in $\theta$ and \elpiIn{equal ~$t'$~ ~$t$~}. Moreover, since \eqfo 
represents the equality of the OL, and since the OL accept term equality up to
\etabeta, then \eqfo also quotient over these reductions.

The \eqho predicate tests if two terms are equal in the world of the meta
languages. This equality function is not capable to understand the \etabeta
reductions over terms of the object language. The ML sees a term of the type
\elpiIn{tm} as atom on which only structural equality can be performed. It is
only possible to dereference a variable if assigned.

\subsection[Compilation: fo\_tm to tm]{Compilation: \elpiIn{fo_tm} $\to$ \elpiIn{tm}}

- how we transform an fo\_tm in tm
- the role of links

\subsection{Unification in ML}

\subsection{Term de-compilation}
% Compile FO to HO
